Verification of the CVM algorithm with a Functional Probabilistic Invariant

### 2025-06-30 paper/sec/related_work.tex
bib fix
Paper: Add reference to paper by Haslbeck,Nipkow 2016 (mentioned by reviewers).

\subsection{Probabilistic Invariants and Formalization}
As far as we know, probabilistic invariants have not been used to establish exponentially decreasing tail-bounds.
However, it is fairly common to use recursive analysis techniques to establish results about expectations or variance of random variables, such as their run-time~\cite[Section 1.4]{motwani1995}.
This is easy due to the linearity of expectations and---for independent random variables---linearity of variances.
A simple example is the Morris counter~\cite{morris1978} or the expected run-time of the quick-sort algorithm~\cite[Section 2.5]{mitzenmacher2005}.

@@

Finally, verification of randomized algorithms expressed in functional programming languages has been tackled by various authors using various proof assistants~\cite{audebaud2009,bosshard2024,demedeiros2024, eberl2020,gopinathan20,hurd03, Probabilistic_Prime_Tests-AFP, tan2024, tassarotti2021}; the most closely related efforts are our mechanizations of frequency moments algorithms~\cite{karayel2022, karayel2023}.
Moreover, for classical imperative randomized algorithms, there are approaches based on probabilistic Hoare logic~\cite{denhartog2002}.
An interesting related work by Haselwarter et al.~\cite{haselwarter2025} discusses some of the issues we encountered in the transformation-based proof (Section~\ref{sec:transformation_based_proof}), such as (approximate) equivalence between related randomized algorithms.
However, because our work reasons directly over the semantics of randomized algorithms expressed in the Giry monad, we did not deeply explore probabilistic logic-based methods.

>>>>

\subsection{Probabilistic Invariants and Formalization}
As far as we know, probabilistic invariants have not been used to establish exponentially decreasing tail-bounds.
However, it is fairly common to use recursive analysis techniques to establish results about expectations or variance of random variables, such as their run-time~\cite[Section 1.4]{motwani1995}.
This is easy due to the linearity of expectations and---for independent random variables---linearity of variances.
A simple example is the Morris counter~\cite{morris1978} or the expected run-time of the quick-sort algorithm~\cite[Section 2.5]{mitzenmacher2017}.

@@

Finally, verification of randomized algorithms expressed in functional programming languages has been tackled by various authors using various proof assistants~\cite{audebaud2009,bosshard2024,demedeiros2024, eberl2020,gopinathan20,haslbeck2016,hurd03,Probabilistic_Prime_Tests-AFP,tan2024,tassarotti2021}; the most closely related efforts are our mechanizations of frequency moments algorithms~\cite{karayel2022, karayel2023}.
Moreover, for classical imperative randomized algorithms, there are approaches based on probabilistic Hoare logic~\cite{denhartog2002}.
An interesting related work by Haselwarter et al.~\cite{haselwarter2025} discusses some of the issues we encountered in the transformation-based proof (Section~\ref{sec:transformation_based_proof}), such as (approximate) equivalence between related randomized algorithms.
However, because our work reasons directly over the semantics of randomized algorithms expressed in the Giry monad, we did not deeply explore probabilistic logic-based methods.
